\chapter{Emergence of Observable Laws} \label{chapter-observable-laws}

Observable physics arises in SCM when symbolic structures—governed purely by return logic—stabilize into persistent, interaction-ready identities. The conservation laws, quantization, and relational observables of physical theory all emerge from the constraints of coherence preservation under RuleEvolution.

This chapter formalizes the emergence of classical laws, quantum behavior, and measurement phenomena as necessary consequences of symbolic reuse logic.

\section{Observables as Coherence Invariants} \label{sec:observables}

Let $[A] \in \Omega_3$ be observable.

\begin{itemize}
  \item \textbf{Mass:} $m([A]) := \Lambda([A])$ (latency-weighted effort),
  \item \textbf{Charge:} $q([A]) := X_\pi([A])$ (return asymmetry),
  \item \textbf{Spin:} symmetry multiplicity in return paths,
  \item \textbf{Position:} reuse anchoring $\mathcal{P}([A])$,
  \item \textbf{Time:} reuse interval $T([A])$.
\end{itemize}

These are not external quantities—they are coherence-based structural invariants.

\section{Conservation Laws as Closure Constraints} \label{sec:conservation}

Let $\mathcal{L}([C])$ be a composite identity involving $\{[A_i]\}$.

\begin{itemize}
  \item \textbf{Mass Conservation:}
  \[
  \sum_i \Lambda([A_i]) = \Lambda([C])
  \]
  \item \textbf{Charge Conservation:}
  \[
  \sum_i X_\pi([A_i]) = X_\pi([C])
  \]
  \item \textbf{Return Closure:}
  \[
  \partial\chi([C]) = 0 \Rightarrow \text{conservation holds}
  \]
\end{itemize}

Noether’s theorem is reinterpreted: every conservation law arises from symbolic return closure, not spacetime symmetry.

\section{Symmetries as Return Redundancies} \label{sec:symmetries}

A system exhibits a symmetry when:

\begin{itemize}
  \item Multiple return paths $\mathcal{L}_i$ lead to the same identity $[A]$,
  \item These paths have equivalent effort:
  \[
  \Lambda(\mathcal{L}_1) = \Lambda(\mathcal{L}_2)
  \]
  \item RuleEvolution preserves both under $\mathcal{F} - \lambda \cdot \mathcal{C}$ optimization.
\end{itemize}

Gauge invariance becomes coherence redundancy: transformations that do not change symbolic persistence.

\section{Collapse and Measurement Revisited} \label{sec:measurement-collapse}

Let observer identity $[O]$ reuse target $[A]$ such that:

\begin{itemize}
  \item $[A] \in \mathcal{L}([O])$,
  \item $\rho([O]) \gg 1$, $\partial\chi([O]) = 0$,
  \item $\rho([A]) \ll 1$, $\partial\chi([A]) \ne 0$.
\end{itemize}

Then $[A]$ undergoes measurement collapse:
\[
\chi([A]) \to \chi'([A]) \in \text{coherence-locked subset}
\]

Measurement is symbolic absorption of return structure by a higher-coherence anchor.

\section{Uncertainty as Reuse Degeneracy} \label{sec:uncertainty}

Let an identity $[A]$ have:

\begin{itemize}
  \item Fixed position: $\mathcal{P}([A])$,
  \item Variable return paths: high $H([A])$,
  \item Nonzero drift: $\partial\chi([A]) \ne 0$.
\end{itemize}

Then:
- The identity’s momentum (drift exposure) is uncertain.
- Exact coherence implies degenerate reuse, not fixed observables.

This is symbolic uncertainty: tradeoff between reuse localization and return redundancy.

\section{Reality as Structural Persistence} \label{sec:reality}

Observable reality is the coherence front:

\[
\mathcal{R}_{\text{real}} := \{ [A] \in \Omega_3 \mid \partial\chi = 0,\ \rho \gg 1,\ \text{interacts without collapse} \}
\]

\begin{itemize}
  \item It is not fixed—it evolves under RuleEvolution,
  \item It is the set of structures that survive symbolic contradiction,
  \item It contains all physical observables by construction.
\end{itemize}

\section{Summary: Laws as Return-Constrained Dynamics} \label{sec:summary-laws}

In SCM, physical laws emerge because:

\begin{itemize}
  \item Return must be closed,
  \item Reuse must be contradiction-minimizing,
  \item Drift must be bounded under interaction,
  \item Latency and entropy must balance for survival.
\end{itemize}

There are no fields or forces—only symbolic structure resolved under RuleEvolution.

