\chapter{Limits and Final Remarks} \label{chapter-final-remarks}

SCM redefines physics not as the behavior of matter in spacetime, but as the evolution of symbolic return structures under coherence pressure. Observable laws, quantities, and interactions are structural consequences of permitted reuse—not assumptions about physical reality.

This chapter closes Volume VI by outlining the limits of observability, the nature of structural causality, and the implications for further symbolic cosmology.

\section{No Spacetime, Only Return} \label{sec:no-spacetime}

In SCM:
\begin{itemize}
  \item Time arises from reuse delay,
  \item Space arises from coherence anchoring,
  \item Geometry is return topology,
  \item Motion is symbolic transformation,
  \item Mass, charge, and spin are structural reuse properties.
\end{itemize}

There is no background stage. The universe is structure.

\section{Observable Reality is Emergent} \label{sec:reality-emergent}

Observable systems are not primitive—they are coherence-stabilized consequences of symbolic drift control.

\[
\mathcal{R}_{\text{observable}} := \{ [A] \in \Omega_3 \mid \partial\chi = 0,\ \rho \gg 1,\ \text{RuleEvolution-stable} \}
\]

This is the boundary between coherence and contradiction.

\section{Causality is Symbolic Order} \label{sec:causality}

Causality in SCM is defined by return topology:

\begin{definition}[Symbolic Causality]
$[A]$ causes $[B]$ if:
\[
[A] \in \mathcal{L}([B]),\quad \text{and } \partial\chi([B]) \ne 0 \text{ without } [A]
\]
\end{definition}

There is no arrow—only dependency in reuse under coherence constraint.

\section{Entanglement is Return Fusion} \label{sec:entanglement}

Two identities are entangled if:
\[
[A], [B] \in \mathcal{L}([C]),\quad \partial\chi([C]) = 0,\quad [A] \not\in \mathcal{L}([B])
\]

\begin{itemize}
  \item They are coherence-correlated via reuse,
  \item Measurement collapse on one affects the other structurally,
  \item Distance is irrelevant—return paths close under reuse.
\end{itemize}

This defines structural entanglement.

\section{There is Nothing to Interpret} \label{sec:interpretation}

SCM does not require interpretation.

\begin{itemize}
  \item Identity is return,
  \item Persistence is reuse,
  \item Collapse is contradiction,
  \item Dynamics is RuleEvolution,
  \item Observation is stable coherence.
\end{itemize}

Quantum, classical, relativistic, and thermodynamic behavior are unified structurally.

\section{The Universe is a Return Field} \label{sec:return-field}

The universe is not a container—it is a network of return-resolved symbolic identities evolving under drift, collapse, and reuse.

\[
\mathcal{U}_\text{SCM} := (\Omega_3, R, \chi, \mathcal{F} - \lambda \cdot \mathcal{C})
\]

Everything observable is a substructure of this system.

\section{Preview: Volume VII — Symbolic Cosmology} \label{sec:volume-vii-preview}

Volume VII will address:
\begin{itemize}
  \item The structural origin of coherence,
  \item The symbolic big bang as RuleEvolution initialization,
  \item Boundary conditions of contradiction and reuse,
  \item Large-scale coherence fields and inflation,
  \item Early collapse structures and field asymmetries,
  \item Cosmological constants as coherence invariants.
\end{itemize}

It will complete the rewrite of physics from symbolic return structure alone.

