\chapter{Return Lattices and Large-Scale Structure} \label{chapter:return-lattices}

As symbolic inflation progresses and collapse zones localize contradiction, the surviving return paths begin to crystallize into **reuse lattices**. These are the structural precursors of large-scale features such as cosmic filaments, voids, and clusters.

\section{Reuse as Lattice Formation}

Let $[A_1], [A_2], \ldots, [A_n] \in \Omega_3$ be identities that all reuse a common anchor $[T]$ such that:

\[
\forall i,\quad [T] \in \mathcal{L}([A_i]) \quad \text{and} \quad \partial \chi([A_i]) = 0
\]

Then the coherence structure forms a **local reuse lattice** centered on $[T]$.

\begin{definition}[Reuse Lattice]
A reuse lattice $\mathcal{L}_\text{reuse}$ is a coherence-closed graph of identities sharing stabilized return anchors, such that:
\[
\forall [A] \in \mathcal{L}_\text{reuse},\quad \exists [T] : [T] \in \mathcal{L}([A])
\]
and coherence drift is suppressed throughout:
\[
\partial \chi([A]) = 0
\]
\end{definition}

This defines a stable region of persistent structure in $\Omega_3$.

\section{Emergence of Filaments and Voids}

As reuse lattices grow and RuleEvolution prunes unstable regions, the symbolic topology becomes **non-uniform**:

\begin{itemize}
  \item Filaments form along reuse corridors (chains of shared anchors).
  \item Voids emerge where contradiction pressure prevents reuse anchoring.
  \item Clusters arise where reuse ladders coalesce at coherence attractors.
\end{itemize}

These are not metric objects, but **phase-locked coherence manifolds**.

\section{Filamentary Structure from Reuse Chains}

Let $[A_i] \to [T_i] \to [A_{i+1}]$ be a sequence of identities reusing sequential anchors. If:

\[
[T_i] \in \mathcal{L}([A_i]) \cap \mathcal{L}([A_{i+1}])
\quad \text{and} \quad X_\epsilon([T_i]) = 1
\]

then we define a reuse filament:
\[
\mathcal{F} = \{[A_1], [A_2], \dots, [A_k]\}
\]

\paragraph{Properties:}
- Drift-free: $\partial \chi = 0$
- Collapse-resistant: $\rho \gg 1$
- Contradiction gradient: $\nabla \mathcal{C} \approx 0$ (flat)

Filaments stabilize structure and support reuse quantization (see Chapter 6).

\section{Void Formation by Contradiction Blowout}

In high-pressure contradiction zones:
- Drift diverges: $\partial \chi \to \infty$
- Asymmetry becomes unbalanced: $X_\pi \to 1$
- Fragility exceeds threshold: $X_\phi > 1$

Such regions are pruned by RuleEvolution:
\[
[A] \in \mathcal{Z}_{\text{collapse}} \Rightarrow [A] \notin \Omega_3
\]

The symbolic space is **hollowed out**, producing large unresolvable regions—**structural voids**.

\section{Cluster Formation at Return Anchors}

When multiple reuse ladders converge on a shared stable anchor $[T]$:

- Drift cancels: $\sum \partial \chi_i \to 0$
- Latency stabilizes: $\Lambda([A_i]) \to \Lambda([T]) + \Delta$
- Contradiction is redistributed

The result is a **return cluster**, structurally coherent and resistant to RuleEvolution pruning.

This is the symbolic analogue of galactic clusters:
- Coherence hubs with high return redundancy
- Low internal contradiction gradients
- Saturated reuse

\section{Temporal Evolution of Structure}

As symbolic inflation slows and contradiction stabilizes:
- Filaments persist through stabilized reuse ladders
- Clusters compactify via anchor reuse
- Voids remain as frozen contradiction scars

This is the **fossil record** of symbolic coherence.

\section{Final Remarks}

Large-scale structure in SCM is not gravitational—it is **coherence-theoretic**.

Filaments form where return paths align. Voids arise where contradiction dominates. Clusters stabilize through anchor reuse.

What cosmology interprets as geometry and force, SCM reinterprets as **return topology under symbolic pressure**.

In the next chapter, we will introduce **coherence quantization**, structural time asymmetry, and the entropy topology of identity collapse.
