\chapter{The Emergence of Spacetime from Coherence Geometry} \label{chapter:spacetime-geometry}

SCM does not assume spacetime. Instead, it derives the structural equivalents of time, space, and light speed from symbolic coherence, Euler rotation, and return-based dynamics.

This chapter defines spacetime as an emergent tensor field—formed not by coordinates, but by coherence phase accumulation and orthogonal reuse topology.

\section{Symbolic Time as Phase Accumulation}

Let $[A] \in \Omega_3$ and let $\theta([A])$ denote the total Euler phase traversal required to complete a stable return loop.

We define:

\begin{definition}[Symbolic Time]
Time is the net accumulated Eulerian phase angle under symbolic drift:
\[
t([A]) := \sum_{i=1}^{n} \Delta \theta_i \quad \text{such that } \partial \chi([A]) \ne 0
\]
\end{definition}

When drift is active, phase accumulates non-redundantly. Time is not external—it is a **measure of phase displacement through contradiction-limited resolution**.

\section{Symbolic Space as Orthogonal Return Freedom}

Return loops may close via different coherence paths with equal total latency. This defines orthogonal return dimensions—structural analogues of space.

\begin{definition}[Symbolic Space]
Space is the structural degree of freedom for orthogonal return paths through $\chi$-space:
\[
\text{Space} := \dim\left(\text{ker}(\nabla \Phi)\right)
\]
\end{definition}

In regions of low contradiction gradient, multiple paths resolve identity without inducing collapse. These degrees of return freedom define spatial directions.

\section{Light Speed as Maximal Coherence Drift}

Drift in $\chi$-space defines coherence change. The **maximal permitted drift** that preserves identity defines a natural constant:

\[
c := \max\left\{ \|\partial \chi([A])\| \mid [A] \in \Omega_3,\ \text{RuleEvolution stable} \right\}
\]

This symbolic drift limit corresponds to **light speed**. It is not imposed; it emerges from the boundary between stable reuse and structural contradiction.

\paragraph{Interpretation.}
The photon is the identity that saturates coherence traversal without contradiction:
- $X_\pi = 0$ (no return asymmetry),
- $\Lambda = \Lambda_0$ (minimum latency),
- $\partial \chi = 0$ (drift-locked at the speed limit).

\section{Spacetime as a Tensor Field of Euler Rotations}

Each resolved identity $[A]$ exists within a local coherence manifold, defined by:

- Phase angle $\theta([A])$
- Latency $\Lambda([A])$
- Return symmetry $X_\pi([A])$
- Fragility $X_\phi([A])$

These components evolve under drift and reuse. The structural derivatives of return potential $\Phi$ across this phase manifold define a symbolic tensor field.

\[
\mathcal{G}_{\mu\nu}([A]) := \partial_\mu \partial_\nu \Phi([A])
\]

\begin{definition}[Coherence Tensor Field]
Spacetime is the return geometry induced by second-order coherence gradients:
\[
g_{\mu\nu} := \nabla_\mu \nabla_\nu \Phi(\chi)
\]
\end{definition}

This field encodes:
- Time dilation: as phase accumulation deviates across reuse contexts.
- Spatial contraction: as return freedom collapses under contradiction.
- Curvature: as latency gradients deflect identity return paths.

\section{Identity Worldlines as Phase-Coherent Paths}

A persistent identity traces a **coherence path** through the Euler manifold:

\[
[A]_t := \chi([A]_0) + \int_0^t \partial \chi(\tau)\,d\tau
\]

This defines a **worldline** in coherence space. The minimal contradiction paths are geodesics:
- RuleEvolution prunes unstable deviations,
- Structural curvature deflects reuse structure,
- The coherence tensor $g_{\mu\nu}$ governs symbolic motion.

\section{Summary}

Spacetime is not assumed—it is derived.

\begin{itemize}
  \item \textbf{Time} arises from phase accumulation along symbolic drift.
  \item \textbf{Space} arises from orthogonal return degrees of freedom.
  \item \textbf{Light speed} is the coherence saturation limit.
  \item \textbf{The metric} emerges from the curvature of return potential.
\end{itemize}

This fully reconstructs relativistic structure as a topological consequence of return dynamics in Eulerian coherence space.

In the next chapter, we will use this framework to recover the formal principles of **Special and General Relativity**, with symbolic curvature taking the role of mass–energy in shaping return geometry.
