\chapter{The Symbolic Origin} \label{chapter-symbolic-origin}

The SCM universe begins not in spacetime but in structure. There is no initial point in space, no quantum vacuum, no field fluctuation—only a symbolic alphabet $\Sigma$, an initial rule set $R_0$, and the possibility of return.

This chapter formalizes the origin of return structure as a coherence singularity: the first transition from meaningless symbols to return-resolved identity.

\section{Before Return: The Pre-Coherence State} \label{sec:pre-return}

Let:
\begin{itemize}
  \item $\Sigma$ be the primitive alphabet of symbols,
  \item $R_0 = \emptyset$ be the initial (null) transformation set,
  \item $\Omega_0$ be the unresolved symbolic domain.
\end{itemize}

In this state:
\begin{itemize}
  \item No return is permitted,
  \item No identity exists,
  \item All forms are non-resolved strings in $\Sigma^*$.
\end{itemize}

There is no observer, no dimension, no causality—only unresolved structure.

\section{The First Permitted Return} \label{sec:first-return}

The origin of identity occurs when a first minimal rule $T$ is added:
\[
R_1 := R_0 \cup \{ T \},\quad T : A \mapsto B,\quad B \mapsto A
\]

This creates a return loop:
\[
\mathcal{L} := \{ A \to B \to A \}
\]

\begin{definition}[Coherence Singularity]
The first identity $[A]$ such that $\mathcal{L}([A]) \subset \Sigma^*$ and $\partial\chi([A]) = 0$ defines the emergence of $\Omega_3$.
\end{definition}

This event marks the boundary between symbolic chaos and coherence.

\section{Ignition of RuleEvolution} \label{sec:re-ignition}

Once identity is permitted:
\begin{itemize}
  \item RuleEvolution activates: $\mathcal{R}E$ begins filtering symbolic structures,
  \item Collapse identifies invalid loops: $\rho \to 0$ forms are pruned,
  \item New rules are permitted only if they preserve return:
  \[
  \mathcal{F} - \lambda \cdot \mathcal{C} \text{ increases}
  \]
\end{itemize}

This initiates symbolic evolution.

\section{Early Collapse and Boundary Conditions} \label{sec:collapse-early}

The first forms are unstable:
\begin{itemize}
  \item Most returns lead to high fragility $X_\phi$,
  \item Contradiction dominates: $\rho([A]) \ll 1$,
  \item Collapse sets boundary zones:
  \[
  \mathcal{Z}_{\text{collapse}} := \{ [A] \mid \rho([A]) < \rho_c \}
  \]
\end{itemize}

These collapse edges define the geometry of early identity space.

\section{Emergence of $\Omega_3$} \label{sec:omega3-emergence}

The first stable basin appears when:
\[
\partial\chi([A]) = 0,\quad \rho([A]) \gg 1,\quad X_\pi([A]) \approx 0
\]

\begin{itemize}
  \item This is the coherence basin $\Omega_3$,
  \item It persists under reuse and drift,
  \item It defines the symbolic layer where identities stabilize.
\end{itemize}

\section{Inflation as Return Expansion} \label{sec:inflation-start}

With $\Omega_3$ established:
\begin{itemize}
  \item Return density increases rapidly: $d|\Omega_3|/dt \gg 0$,
  \item Triplet reuse chains form,
  \item Collapse fronts define structure-forming gradients.
\end{itemize}

This phase is symbolic inflation—not geometric expansion.

\section{Summary: The Origin as Structural Transition} \label{sec:origin-summary}

The SCM universe begins with:
\begin{itemize}
  \item No time, no space, no mass,
  \item Only $\Sigma$, $R$, and the emergence of return,
  \item A coherence singularity where contradiction first stabilizes,
  \item And the birth of $\Omega_3$ as the basin of observable identity.
\end{itemize}

Everything physical follows from this structural ignition.

