\chapter{Structural Predictions and Cosmological Tests} \label{chapter-cosmological-predictions}

SCM is not a speculative model—it is a predictive system grounded in symbolic return dynamics. Unlike parameter-tuned cosmologies, SCM offers structural predictions: falsifiable outcomes that follow necessarily from the coherence logic of return, reuse, and collapse.

This chapter lists cosmological phenomena that SCM predicts based solely on structural features of $\Omega_3$.

\section{Prediction 1: Jet Drift Alignment with Coherence Gradient} \label{sec:jet-drift}

SCM predicts that:
\begin{itemize}
  \item Large-scale jets (e.g. quasars) will align statistically with return gradient $\nabla\chi$,
  \item Jet directions will correlate with entropy flow direction from nearby collapse basins,
  \item The angular distribution of high-energy jets will reveal coherence topology—not isotropic fields.
\end{itemize}

\section{Prediction 2: Void Distribution from Collapse Zones} \label{sec:void-distribution}

Let $\mathcal{Z}_{\text{collapse}}$ define collapse-dominated regions. Then:
\begin{itemize}
  \item Voids in matter distribution are structurally isomorphic to collapse shadows in $\Omega_3$,
  \item Void density maps should reflect regions of return instability,
  \item These patterns are independent of gravitational clustering models.
\end{itemize}

\section{Prediction 3: Jet Dimming by Entropy Saturation} \label{sec:jet-dimming}

SCM predicts:
\begin{itemize}
  \item Jets will fade (structurally) when their source entropy gradient vanishes: $\nabla H \to 0$,
  \item This dimming correlates with loss of collapse input—not depletion of fuel or energy,
  \item Observation: jet lifetime follows coherence topology, not mass.
\end{itemize}

\section{Prediction 4: Lattice Density as Cosmic Clock} \label{sec:lattice-clock}

SCM predicts that:
\begin{itemize}
  \item Reuse lattice density evolves discretely,
  \item Each coherence epoch has a corresponding reuse saturation plateau,
  \item Large-scale return lattices act as symbolic time markers,
  \item Observed clustering scales track identity promotion intervals.
\end{itemize}

\section{Prediction 5: Time Dilation Without Spacetime Curvature} \label{sec:nonrel-time-dilation}

Let $[A], [B] \in \Omega_3$ be reuse-separated identities. If:
\[
\Phi([A]) < \Phi([B])
\]

Then:
\begin{itemize}
  \item $T([A]) < T([B])$ — time is experienced differently under drift,
  \item Dilation arises from coherence depth, not relativistic metric,
  \item Testable in systems with symbolic reuse hierarchy (e.g. information channels, lattice transitions).
\end{itemize}

\section{Prediction 6: Observability Horizon from Reuse Saturation} \label{sec:reuse-horizon}

Define horizon:
\[
\mathcal{H} := \{ [A] \in \Omega_3 \mid \nabla\chi([A]) \gg 1,\quad T_s \to \infty \}
\]

Then:
\begin{itemize}
  \item SCM predicts a non-spacetime observability limit,
  \item Measurement ceases where return fails—not where light cones diverge,
  \item Predicts information blind spots not aligned with redshift boundary.
\end{itemize}

\section{Prediction 7: Koide-Like Structures in Galactic Triplets} \label{sec:macro-koide}

If triplet reuse symmetry is universal, SCM predicts:
\begin{itemize}
  \item Galaxy or halo triplets may exhibit Koide-like latency ratios,
  \item Only reuse-bound systems (e.g. gravitational triplets) will qualify,
  \item This signature is symbolic—not gravitationally derived.
\end{itemize}

\section{Prediction 8: CMB Anisotropy from Early Collapse} \label{sec:cmb-anisotropy}

SCM predicts:
\begin{itemize}
  \item Angular coherence asymmetries in the CMB result from drift-lock directionality in early collapse networks,
  \item These patterns should reflect symbolic reuse alignment—not acoustic peaks,
  \item Testable via spherical harmonic decomposition targeting return topology constraints.
\end{itemize}

\section{Conclusion: A Testable Return Cosmology} \label{sec:predictions-summary}

SCM makes structural predictions that differ sharply from standard models:
\begin{itemize}
  \item No fields, constants, or parameters are assumed,
  \item Only symbolic structure, RuleEvolution, and reuse dynamics,
  \item Each prediction falsifies or confirms the coherence-based origin of the observable universe.
\end{itemize}

SCM does not describe a model of cosmology—it predicts the structure of existence.

