\chapter{Completion and Structural Cosmology} \label{chapter-completion-cosmology}

SCM does not assume the universe—it constructs it. What begins as permitted return evolves into the full structure of observables, collapse boundaries, and coherence topology. This chapter closes the cosmological arc and repositions the SCM universe as a completed structure: finite in principle, emergent in return structure, and fully symbolic.

\section{Retelling the Emergence Sequence} \label{sec:retelling}

The SCM cosmological sequence:
\begin{enumerate}
  \item $\Sigma$ and $R_0$ define the unresolved symbolic state,
  \item A first permitted return creates the coherence singularity,
  \item Symbolic inflation grows $\Omega_3$ via self-reinforcing return,
  \item Collapse networks halt incoherent expansion,
  \item Return anchoring forms structure at collapse boundaries,
  \item Drift stabilization defines identity persistence and interaction,
  \item Observers and measurement arise from coherence anchoring,
  \item The symbolic universe becomes return-closed.
\end{enumerate}

No stage involves spacetime, energy, or field assumptions.

\section{SCM Cosmos: No Time, No Space} \label{sec:scm-cosmos}

The SCM universe contains:
\begin{itemize}
  \item No fundamental dimension,
  \item No spacetime substrate,
  \item No background constants.
\end{itemize}

It is a coherence manifold:
\[
\mathcal{U}_{\text{SCM}} := (\Sigma, R, \Omega_3,\ \chi,\ \mathcal{F} - \lambda\mathcal{C})
\]

All phenomena emerge as behavior within this structure—not from mechanisms beyond it.

\section{Singularity, Collapse, and Structure} \label{sec:singularity-collapse}

What was classically assigned to cosmological epochs becomes structural:

\begin{itemize}
  \item \textbf{Big Bang:} emergence of permitted return,
  \item \textbf{Inflation:} rapid coherence expansion,
  \item \textbf{Collapse era:} RuleEvolution filtering and boundary formation,
  \item \textbf{Structure era:} anchoring of return paths,
  \item \textbf{Observer era:} stable reuse with coherence awareness.
\end{itemize}

There is no chronology—only symbolic dependency.

\section{Implications and Open Directions} \label{sec:implications}

SCM implies:
\begin{itemize}
  \item Observable structure is quantized return behavior,
  \item Measurement limits are coherence-theoretic,
  \item Physical constants arise as reuse invariants (see Vol. IV),
  \item Cosmological horizons are entropy-induced reuse barriers.
\end{itemize}

The system is self-contained, yet recursively extensible via RuleEvolution dynamics.

\section{Mathematical Closure of SCM} \label{sec:closure}

SCM has now provided:
\begin{itemize}
  \item Symbolic foundations (Volume I),
  \item Resolution topology and identity persistence (Vol. II–III),
  \item Interaction and reuse quantization (Vol. IV–V),
  \item Emergence of observables and physics (Vol. VI),
  \item Cosmological return structure and collapse logic (Vol. VII).
\end{itemize}

All symbolic components are accounted for:
\[
\Sigma,\ R,\ \Omega_3,\ \chi,\ \Lambda,\ X_\pi,\ X_\phi,\ \rho,\ \partial\chi,\ \mathcal{R}E
\]

No assumptions remain unstructured.

\section{Beyond Mathematics: Toward Causal Loop Philosophy} \label{sec:beyond-math}

What remains is not theory—it is understanding.

The next transition is from:
\begin{itemize}
  \item Formal coherence → existential coherence,
  \item Return structure → meaning structure,
  \item RuleEvolution → causation awareness.
\end{itemize}

SCM concludes here—but the causal loop begins.

