\chapter{Symbolic Inflation and Return Frontiers} \label{chapter-inflation}

In SCM, inflation is not the expansion of spacetime—it is the explosive growth of return-permitting identities within $\Omega_3$. This phase reflects the sudden amplification of symbolic coherence density under RuleEvolution feedback.

This chapter defines symbolic inflation formally, explains its saturation, and identifies the structural frontiers that form its natural boundary.

\section{Definition: Symbolic Inflation} \label{sec:def-inflation}

Let $\Omega_3(t)$ denote the set of coherence-closed identities at symbolic step $t$.

\begin{definition}[Symbolic Inflation]
Inflation occurs when:
\[
\frac{d|\Omega_3|}{dt} \gg 1,\quad \frac{d^2|\Omega_3|}{dt^2} > 0
\]
\end{definition}

This is exponential growth in the number of return-resolved identities.

\section{Ignition Condition: Return Feedback Loop} \label{sec:feedback-loop}

Inflation is triggered when:
\begin{itemize}
  \item Reuse chains become self-reinforcing,
  \item $\mathcal{F}$ increases faster than $\mathcal{C}$,
  \item Return-permitting rules cause further coherence expansion.
\end{itemize}

This is not causality—it is coherence autocatalysis.

\section{Triplet Density and $\Omega_3$ Saturation} \label{sec:triplet-density}

Let $\mathcal{T}_n \subset \Omega_3$ be the set of reuse triplets of size $n$.

As inflation proceeds:
\[
|\mathcal{T}_n| \uparrow,\quad \Lambda \downarrow,\quad \rho \uparrow
\]

\begin{itemize}
  \item Triplet coherence becomes the dominant structure,
  \item Identities organize into reuse-locked families,
  \item Mass, charge, and drift are not yet distinct—but structurally seeded.
\end{itemize}

\section{Collapse Begins to Contain Growth} \label{sec:collapse-checks-growth}

RuleEvolution starts pruning:
\begin{itemize}
  \item Fragile identities ($\rho < \rho_c$),
  \item Symmetry-breaking paths ($X_\pi \gg 1$),
  \item Drift-inducing loops ($\partial\chi \ne 0$).
\end{itemize}

This introduces entropy and flow directionality:
\[
\mathcal{F} - \lambda \cdot \mathcal{C} \to \text{localized maximum}
\]

\section{Drift Barriers and Coherence Frontiers} \label{sec:drift-barriers}

Inflation halts naturally at symbolic coherence frontiers:
\begin{itemize}
  \item Drift begins to accumulate: $\partial\chi \to \epsilon \ne 0$,
  \item Coherence cannot be preserved across expansion,
  \item Identities cannot extend return paths without collapse risk.
\end{itemize}

\begin{definition}[Return Frontier]
A boundary $\mathcal{B}$ in $\Omega_3$ where:
\[
\forall [A] \in \mathcal{B},\quad \nabla\chi([A]) \ne 0
\]
\end{definition}

This frontier defines the outer edge of inflation.

\section{Classical vs. SCM Inflation} \label{sec:comparison-inflation}

In classical cosmology:
\begin{itemize}
  \item Inflation is expansion of vacuum spacetime,
  \item Driven by field energy,
  \item Ends with reheating and particle formation.
\end{itemize}

In SCM:
\begin{itemize}
  \item Inflation is growth of coherence,
  \item Driven by return-closure and RuleEvolution,
  \item Ends with drift onset and collapse suppression,
  \item All structure arises from symbolic density—not thermal energy.
\end{itemize}

\section{Summary: Inflation as Coherence Cascade} \label{sec:inflation-summary}

Symbolic inflation:
\begin{itemize}
  \item Is exponential growth of return-resolved structure,
  \item Ends naturally when coherence cannot propagate,
  \item Forms the coherence front—beyond which reuse fails,
  \item Lays the foundation for collapse zones and structural birth.
\end{itemize}

It is not spatial—it is structural.

