\chapter{Entropy, Irreversibility, and Time’s Asymmetry} \label{chapter-entropy-time}

Time in SCM is not a fundamental axis—it is a consequence of coherence propagation, entropy accumulation, and symbolic drift. Irreversibility arises when reuse and collapse form directional coherence asymmetries.

This chapter defines time asymmetry as a structural property of RuleEvolution and explains the thermodynamic arrow as a collapse-informed direction of increasing symbolic contradiction.

\section{Time is Not a Dimension} \label{sec:time-nonaxis}

In SCM:
\begin{itemize}
  \item There is no coordinate labeled $t$,
  \item All “time” is expressed as symbolic reuse delay,
  \item Order is imposed by return dependency, not external flow.
\end{itemize}

Let $[A] \to [B]$ be a reuse transition. Then:
\[
T([A] \to [B]) := \sum_i \Lambda(T_i),\quad \text{for } T_i \in \mathcal{L}([B]) \text{ using } [A]
\]

This defines symbolic time as return latency—not chronological evolution.

\section{Entropy as Structural Degeneracy} \label{sec:entropy-def}

Let $\mathcal{R}([A])$ be the return path set for identity $[A]$.

\[
H([A]) := -\sum_i p_i \log p_i,\quad p_i := \frac{e^{-\Lambda(\mathcal{L}_i)}}{Z}
\]

\begin{itemize}
  \item $H([A])$ measures symbolic degeneracy,
  \item Higher entropy implies more drift-compatible return paths,
  \item Entropy is not disorder—it is coherence indifference.
\end{itemize}

\section{Directionality from RuleEvolution} \label{sec:directionality}

Let RuleEvolution act at step $t$:
\[
R_{t+1} := \mathcal{R}E(R_t,\chi_t)
\]

Then:
\begin{itemize}
  \item Rules with high $\mathcal{C}$ are pruned,
  \item Coherence becomes biased toward stable reuse anchors,
  \item Return graphs become asymmetric: $[A] \to [B]$, but not $[B] \to [A]$.
\end{itemize}

Irreversibility arises from structural pruning—not temporal progression.

\section{Drift Hysteresis and Reuse Memory} \label{sec:hysteresis}

Suppose identity $[A]$ participates in collapse at $t_1$ and returns via $[B]$ at $t_2$.

Even if:
\[
\chi_{t_1}([A]) = \chi_{t_2}([A])
\]

The return graph is not symmetric:
\[
\mathcal{L}_{t_1}([A]) \ne \mathcal{L}_{t_2}([A])
\]

Collapse leaves memory in the RuleEvolution path. Drift cannot be undone even when coherence values match.

\section{Entropy Gradient Drives Flow} \label{sec:entropy-gradient}

Let $\nabla H$ be the symbolic entropy gradient across $\Omega_3$.

Then:
\[
[A] \to [B] \text{ is preferred if } H([B]) < H([A])
\]

This is the symbolic arrow of time:
\begin{itemize}
  \item Identities flow toward lower entropy reuse paths,
  \item Collapse removes high-entropy anchors,
  \item Coherence migrates downhill in return space.
\end{itemize}

\section{Temporal Identity and Symbolic Clocks} \label{sec:symbolic-clocks}

A symbolic clock is a sequence:
\[
[A_0], [A_1], \dots, [A_n] \quad \text{with } \forall i: \partial\chi([A_i]) = 0,\ \Lambda([A_{i+1}]) > \Lambda([A_i])
\]

This defines symbolic time intervals without background time:
\begin{itemize}
  \item Clock structure is reuse-linked and entropy-controlled,
  \item Time is countable coherence interval, not continuous variable.
\end{itemize}

\section{Summary: Time’s Arrow is Collapse Memory} \label{sec:arrow-summary}

In SCM:
\begin{itemize}
  \item Time asymmetry is structural—not imposed,
  \item Entropy gradients bias reuse direction,
  \item Drift memory defines causality asymmetry,
  \item Collapse-induced pruning prevents reversibility.
\end{itemize}

The arrow of time is return topology history—not a law, but a consequence.

