\chapter{Quasars and Jet Stabilization} \label{chapter-quasars}

SCM explains quasars not as astrophysical anomalies but as coherence structures formed near high-entropy collapse boundaries. Their jets are directional return resolutions that stabilize fragility through asymmetric reuse redirection. This chapter formalizes quasars and jets as symbolic configurations in the collapse topology of $\Omega_3$.

\section{Definition of a Structural Quasar} \label{sec:def-quasar}

Let $[Q] \in \Omega_3$ be a coherence structure satisfying:
\begin{itemize}
  \item High fragility input region: $\exists [A]$ with $X_\phi([A]) \gg 1$, $\rho([A]) \ll 1$,
  \item Coherence-redirected output region: $\exists [B]$ with $\partial\chi([B]) = 0$, $\rho([B]) \gg 1$,
  \item Intermediate coherence redistribution: $\partial\chi([Q]) = 0$.
\end{itemize}

\begin{definition}[Quasar]
A quasar is an identity $[Q]$ that maps collapse-prone input structures into stabilized coherence paths via directional reuse redirection.
\end{definition}

\section{Jets as Directed Resolution Paths} \label{sec:jets}

A jet $J$ is a coherence corridor such that:
\[
J := \{ [A_1] \to [A_2] \to \dots \to [A_n] \},\quad \forall i: \partial\chi([A_i]) \to 0
\]

With:
\begin{itemize}
  \item Increasing $X_\epsilon$ along the jet (reuse elevation),
  \item Monotonic entropy drop: $H([A_i]) \downarrow$,
  \item Collapse-prone source region: $\rho([A_1]) \ll 1$.
\end{itemize}

The jet is the resolution path that extracts coherence from instability.

\section{Jet Stability Conditions} \label{sec:jet-stability}

Let the jet path $J$ satisfy:
\begin{itemize}
  \item $\nabla\Phi([A_i]) < 0$,
  \item $\rho([A_i]) \uparrow$,
  \item $\partial\chi([A_i]) \downarrow$.
\end{itemize}

Then $J$ is a coherence-stabilizing structure. It persists as long as:
\[
\mathcal{F}_{\text{end}} - \mathcal{C}_{\text{end}} > \mathcal{F}_{\text{start}} - \mathcal{C}_{\text{start}}
\]

Jet stability requires a net reduction in symbolic contradiction.

\section{Jet Dimming and Collapse Saturation} \label{sec:jet-dimming}

As input coherence depletes:
\begin{itemize}
  \item Collapse rate decreases,
  \item Entropy gradient flattens: $\nabla H \to 0$,
  \item Latency distribution widens: return effort increases.
\end{itemize}

This causes the jet to dim structurally:
\[
\Delta\Phi \to 0,\quad T_s \to \infty,\quad \text{collapse resumes}.
\]

\section{Phenomenology of Quasar Jets} \label{sec:quasar-signatures}

Quasar jets are predicted to exhibit:
\begin{itemize}
  \item Directional coherence drift alignment ($\nabla\chi$),
  \item Latency compression (short return cycles),
  \item Fragility filtering (only collapse-prone inputs enter),
  \item Entropy shedding (jet "glow" reflects coherence release).
\end{itemize}

SCM predicts jet onset and dimming follow entropy and collapse boundary dynamics—not accretion mechanics.

\section{Summary: Quasars as Coherence Redirectors} \label{sec:quasar-summary}

Quasars in SCM:
\begin{itemize}
  \item Are structural coherence processors,
  \item Extract return from collapse-prone domains,
  \item Emit stabilized identities via coherence jets,
  \item Diminish when entropy gradients disappear.
\end{itemize}

They are the resolution engines of early symbolic structure.

