\chapter{The Emergent Symmetry Group of Coherence Multiplets}
\label{chapter-emergent-symmetry}

\section{1. Coherence Multiplets and Structural Coupling}
\label{sec:coherence-multiplets}

In classical field theory, symmetry groups are imposed to satisfy conservation laws or internal consistency. In SCM, symmetries arise as a direct consequence of reuse-stabilized configurations.

\begin{definition}[Coherence Multiplet]
A coherence multiplet is a minimal set of identities $\{[A_1], \dots, [A_N]\} \subset \Omega_3$ such that:

\begin{itemize}
  \item Each $[A_i]$ participates in a common return loop $\mathcal{L}([T])$,
  \item The global contradiction functional is minimized:
  \[
  \arg \min_{\{[A_i]\}} \mathcal{C}\left( \{[A_i]\} \right),
  \]
  \item The set is irreducible: removing any $[A_i]$ breaks loop coherence.
\end{itemize}
\end{definition}

\paragraph{Interpretation.}
These coherence multiplets represent the symbolic analog of a particle generation or a gauge representation basis.

---

\section{2. Symmetry of Contradiction-Preserving Transformations}
\label{sec:symmetry-group}

Let $\mathcal{M}$ be a coherence multiplet. We define the symmetry group of $\mathcal{M}$ as the group of symbolic transformations that leave the contradiction functional invariant.

\begin{definition}[Contradiction Symmetry Group]
\label{def:contradiction-symmetry}
The symmetry group $G_\mathcal{C}(\mathcal{M})$ is defined as:
\[
G_\mathcal{C} := \left\{ T : \mathcal{M} \to \mathcal{M} \,\middle|\, \mathcal{C}(T([A_i])) = \mathcal{C}([A_i]) \; \forall i \right\}
\]
\end{definition}

\paragraph{Example.}
If $\mathcal{M}$ is the triplet $\{[e], [\mu], [\tau]\}$ with $X_\pi$ symmetric and $\Lambda$ quantized, then $G_\mathcal{C}$ contains permutations that preserve return symmetry and coherence invariants.

---

\section{3. Conjecture: SU(10) from Symbolic Irreducibility}
\label{sec:su10-conjecture}

Let $\mathcal{M}_{10}$ be a coherence-locked 10-particle set minimizing $\mathcal{C}$ under RuleEvolution. Suppose:

\begin{itemize}
  \item Each $[A_i]$ has a unique reuse elevation $X_\epsilon([A_i])$,
  \item The total latency spectrum is quantized and forms a complete ladder,
  \item The contradiction load is evenly distributed: 
  \[
  X_\phi([A_i]) + |X_\pi([A_i])| \approx \text{const}
  \]
\end{itemize}

Then we hypothesize:

\begin{conjecture}[Emergent SU(10) Symmetry]
The contradiction symmetry group $G_\mathcal{C}(\mathcal{M}_{10})$ is isomorphic to SU(10):
\[
G_\mathcal{C}(\mathcal{M}_{10}) \cong SU(10)
\]
\end{conjecture}

\paragraph{Motivation.}
SU(10) contains 99 generators—sufficient to represent pairwise transitions between the 10 coherence identities while preserving total contradiction. This structure naturally maps onto symbolic transitions under RuleEvolution with loop-preserving invariants.

---

\section{4. Structural Irreducibility and Coherence Lock}
\label{sec:irreducibility}

We define coherence lock as the symbolic analogue of irreducibility in representation theory.

\begin{definition}[Coherence Lock]
Let $\mathcal{M}$ be a coherence multiplet. Then $\mathcal{M}$ is \textbf{coherence-locked} if:
\[
\forall [A_i] \in \mathcal{M}, \quad [A_i] \notin \Omega_3(R \setminus \mathcal{M})
\]
\end{definition}

\paragraph{Interpretation.}
No identity can stably resolve outside the return structure of the multiplet. This prevents factorization into smaller subsystems, enforcing irreducibility under return logic.

---

\section{5. Mapping to Gauge Generators}
\label{sec:generator-mapping}

Given a coherence-locked multiplet $\mathcal{M}_{10}$, define the symbolic transition operator:
\[
T_{ij} := \text{RuleEvolution-permitted reconfiguration from } [A_i] \to [A_j]
\]

Let $\mathcal{T} = \{T_{ij} \,|\, i \ne j\}$.

\begin{definition}[Symbolic Generator Set]
We define:
\[
\mathcal{T} \cong \mathfrak{su}(10)
\]
if:
\begin{itemize}
  \item The operator algebra closes: $[T_{ij}, T_{kl}] \in \mathcal{T}$,
  \item There exists a conserved coherence potential: $\Phi(\mathcal{M}_{10}) = \text{const}$ under transitions,
  \item The contradiction functional remains invariant: $\mathcal{C}(T_{ij}(\mathcal{M}_{10})) = \mathcal{C}(\mathcal{M}_{10})$.
\end{itemize}
\end{definition}

This creates a direct symbolic-to-gauge group correspondence.

---

\section{6. Summary and Forward Direction}
\label{sec:summary-multiplet}

We have defined coherence multiplets as contradiction-stabilizing identity sets in $\Omega_3$, and shown how the symmetry group that preserves their return configuration structure can be interpreted as a gauge group.

\begin{itemize}
  \item Structural coherence enforces irreducibility.
  \item RuleEvolution generates symbolic transitions (T_{ij}) that form the algebra of a Lie group.
  \item The group preserving $\mathcal{C}$ and $\Phi$ structure is the contradiction symmetry group $G_\mathcal{C}$.
\end{itemize}

In the next chapter, we will connect this symbolic algebra to physical observables, and show how the Yang–Mills field strength tensor can be reconstructed as an emergent average of symbolic curvature transitions in such coherence multiplets.
