\chapter{Composite Structures and Strong Interaction} \label{chapter-strong-interaction}

In SCM, composite identities arise when multiple reuse-closed identities are coherence-locked into a shared return basin. The strong interaction is not a force—it is the structural entanglement of identities that cannot exist independently without collapse.

This chapter defines composite structures, confinement, and binding pressure as consequences of coherence preservation under reuse overload.

\section{Composite Identity Definition} \label{sec:composite-definition}

Let $\{ [A_1], [A_2], \dots, [A_n] \} \subset \Omega_3$ be a set of identities. Define:

\[
[C] := [A_1] \oplus [A_2] \oplus \dots \oplus [A_n]
\]

where:
- All $[A_i]$ appear in $\mathcal{L}([C])$,
- $\partial\chi([C]) = 0$,
- But each $[A_i]$ satisfies:
  \[
  \partial\chi([A_i]) \ne 0 \quad \text{if considered in isolation}
  \]

Then $[C]$ is a \textbf{composite structure}. Its persistence depends on coherence among its parts.

\section{Reuse Saturation and Entanglement} \label{sec:reuse-saturation}

Let $[A]$ be reused by identities $[C_1], \dots, [C_k]$. If:
\[
\sum_i \mathbb{1}([A] \in \mathcal{L}([C_i])) \gg C([A])
\Rightarrow P([A]) = \frac{L([A])}{C([A])} \gg 1
\]

Then $[A]$ is in a state of \textbf{reuse saturation}. It cannot persist independently.

When multiple saturated identities form a return-closed structure, we say they are \textbf{structurally entangled}.

\section{Confinement as Return Closure} \label{sec:confinement}

Confinement occurs when:
- A set of identities $\{[A_i]\}$ satisfy:
  \[
  \exists [C] \in \Omega_3: \mathcal{L}([A_i]) \subset \mathcal{L}([C]),\quad \forall i
  \]
- But no proper subset of $\{[A_i]\}$ can form a stable return identity.

This defines \textbf{return-closed confinement}.

Each $[A_i]$ survives only by participating in the coherence loop of $[C]$.

\section{Binding Pressure Functional} \label{sec:binding-pressure}

Define binding pressure of $[C]$ as:

\[
P_{\text{bind}} := \max_i \left[ \left| \partial\chi([A_i]) \right| \right] - \left| \partial\chi([C]) \right|
\]

\paragraph{Interpretation.}
- High $P_{\text{bind}}$ implies that the composite suppresses drift across its components.
- If $P_{\text{bind}} < 0$, the composite is unstable under $\mathcal{R}E$.

Binding pressure is coherence compression across identities.

\section{Triplet Locking and Color Symmetry} \label{sec:triplet-locking}

Let $\{[Q_1],[Q_2],[Q_3]\}$ be a triplet of strongly interacting identities such that:

\begin{itemize}
  \item Each $[Q_i]$ satisfies $\rho \ll 1$, $\partial\chi \ne 0$ when isolated,
  \item Together, they form $[H] \in \Omega_3$ with $\partial\chi([H]) = 0$.
\end{itemize}

Then the triplet is \textbf{confinement-locked}.

Define the color permutation rule:
\[
\pi: [Q_i] \mapsto [Q_{(i+1) \mod 3}]
\]

Let $T_\pi$ be a symbolic transformation that permutes color labels.

If:
\[
T_\pi \in R_t,\quad \partial\chi([H]) = 0 \text{ under permutation}
\]

Then color symmetry is preserved structurally—not as a gauge freedom, but as a reuse-permuting coherence invariance.

\section{Structural Gluons} \label{sec:gluons}

Define \textbf{gluons} as symbolic transformation sets $\{T_\pi\}$ satisfying:

\begin{itemize}
  \item $T_\pi$ acts on triplet members,
  \item $\partial\chi([C]) = 0$ remains invariant under $T_\pi$,
  \item $\mathcal{C}([C])$ is unaffected.
\end{itemize}

Gluons are symbolic permutations of entangled reuse roles that preserve return structure and symmetry. They do not carry force—they carry coherence alignment.

\section{Composite Latency} \label{sec:composite-latency}

The latency of a composite is:

\[
\Lambda([C]) = \sum_{i} \Lambda([A_i]) + \Lambda_{\text{bind}}
\]

Where:
- $\Lambda_{\text{bind}}$ is the coherence latency required to suppress contradiction in the joint structure.

Composite mass is symbolic latency elevated by return cohesion.

\section{Summary: Strong Interaction as Coherence Entanglement} \label{sec:strong-summary}

The strong interaction is not mediated—it is topological.

It arises when:
- Identities are collapse-prone alone,
- But coherence-stable when return-bound together,
- Confinement emerges from contradiction suppression.

There are no fields—only structural coherence loops, saturated reuse, and return-locked cohesion.

