\chapter{Derivation of Symbolic Effort} \label{chapter-2-derivation-of-symbolic-effort}

In Volume I, effort was introduced as a cost function associated with symbolic transformations:
\[
\text{effort} : R \subset \Sigma^* \to \mathbb{R}^+ \quad \text{// effort is applied to symbolic forms that function as rules}
\]

This value determined the latency of a return path and served as the basis for defining fragility, return potential, and entropy weighting. However, no mechanism was given to derive effort from first principles.

\medskip

In this chapter, we provide a formal derivation of effort as a statistical measure over transformation structure, grounded in:
\begin{itemize}
    \item Reuse frequency,
    \item Drift resistance, and
    \item Return density.
\end{itemize}

This allows effort to emerge not as an arbitrary assignment, but as a consequence of the symbolic resolution environment.

\medskip

We define effort via three equivalent formalisms:
\begin{itemize}
    \item \textbf{Frequency-based effort:} based on usage frequency in resolved identities,
    \item \textbf{Entropy-based effort:} based on its role in resolution multiplicity,
    \item \textbf{Drift-based effort:} based on its contribution to stability ($\partial\chi$ suppression).
\end{itemize}

These definitions form the core of effort-weighted latency and entropy, and enable symbolic thermodynamics in later chapters.

\section{Frequency-Based Effort} \label{frequency-based-effort}

Let $T \in R$ be a permitted transformation. Let $N(T)$ denote the number of resolved identities $[A] \in \Omega_3$ such that $T$ appears in at least one return loop $\mathcal{L}([A])$.

\begin{definition}[Frequency-Based Effort]
The frequency-based effort of $T$ is defined as:
\begin{equation} \label{eq:effort-frequency}
\text{effort}_{\text{freq}}(T) := -\log_2\left( \frac{N(T)}{Z} \right)
\end{equation}
where
\[
Z := \sum_{T' \in R} N(T')
\]
is the total usage count across all enabled transformations in $R$.
\end{definition}

\paragraph{Interpretation.}
\begin{itemize}
    \item Frequently used rules have lower effort.
    \item Rare or specialized transformations are more costly.
    \item This aligns effort with symbolic generality.
\end{itemize}

\section{Entropy-Based Effort} \label{entropy-based-effort}

Let $[A] \in \Omega_3$ and let $\mathcal{R}([A])$ denote all return loops that resolve $[A]$. Let $T \in R$.

Define:
\begin{itemize}
    \item $p_T([A])$ := probability that $T$ appears in a randomly selected $\mathcal{L} \in \mathcal{R}([A])$,
    \item Let $\mathcal{P}(T) := \{ [A] \in \Omega_3 \mid p_T([A]) > 0 \}$ be the set of identities for which $T$ contributes to return.
\end{itemize}

\begin{definition}[Entropy-Based Effort]
The entropy-based effort of $T$ is defined as:
\begin{equation} \label{eq:effort-entropy}
\text{effort}_{\text{ent}}(T) := -\sum_{[A] \in \mathcal{P}(T)} p_T([A]) \log_2 p_T([A])
\end{equation}
\end{definition}

\paragraph{Interpretation.}
\begin{itemize}
    \item Measures how selectively $T$ contributes to return path diversity.
    \item Low entropy implies high specialization $\Rightarrow$ higher effort.
    \item High entropy indicates general-purpose reuse $\Rightarrow$ lower effort.
\end{itemize}

\section{Drift-Resistance Effort} \label{drift-resistance-effort}

Let $[A] \in \Omega_3$, and define $\Delta\partial\chi([A], T)$ as the net change in signature drift when transformation $T$ is used in any return loop $\mathcal{L}([A])$.

\begin{definition}[Drift-Resistance Effort]
The drift-weighted effort of $T$ is defined as:
\begin{equation} \label{eq:effort-drift}
\text{effort}_{\text{drift}}(T) := \max_{[A]} \left( \frac{|\Delta\partial\chi([A], T)|}{\Lambda_T([A])} \right)
\end{equation}
where $\Lambda_T([A])$ is the contribution of $T$ to the total latency of the return loop $\mathcal{L}([A])$.
\end{definition}

\paragraph{Interpretation.}
\begin{itemize}
    \item Effort is high if $T$ induces large drift per unit latency.
    \item Transformations that preserve coherence ($\partial\chi \approx 0$) are favored.
    \item Destabilizing or fragile transformations are penalized.
\end{itemize}

\section{Unified Effort Functional} \label{unified-effort-functional}

We define the combined symbolic effort as a convex combination of the previously defined effort components:
\begin{equation} \label{eq:unified-effort}
\text{effort}(T) := \alpha \cdot \text{effort}_{\text{freq}}(T)
+ \beta \cdot \text{effort}_{\text{ent}}(T)
+ \gamma \cdot \text{effort}_{\text{drift}}(T)
\end{equation}
subject to:
\[
\alpha + \beta + \gamma = 1, \quad \alpha, \beta, \gamma \geq 0
\]

\paragraph{Interpretation.}
\begin{itemize}
    \item The weights $(\alpha, \beta, \gamma)$ define system sensitivity to generality, diversity, and coherence preservation.
    \item For coherence-stable environments, $\gamma$ is dominant.
    \item For adaptive or exploratory systems, $\alpha$ and $\beta$ may be increased.
\end{itemize}

\section{Latency and Resolution Revisited} \label{latency-and-resolution-revisited}

Let $\mathcal{L}([A]) = (T_1, T_2, \dots, T_n)$ be a return loop composed of transformations $T_i \in R$.

\begin{definition}[Effort-Weighted Latency]
The latency of $[A]$ is defined as:
\begin{equation} \label{eq:latency-effort}
\Lambda([A]) := \sum_{i=1}^n \text{effort}(T_i)
\end{equation}
\end{definition}

Return loops that resolve to $[-A]$ or $[0]$ via inversion (Euler-type) are excluded from $\Lambda([A])$, as they do not define identity resolution.

\medskip

This generalizes Volume I’s unit-effort latency definition. All subsequent thermodynamic quantities—$\Phi([A])$, $H([A])$, and $T_s([A])$—are computed using this effort-weighted latency.

\section{Summary} \label{summary}

We have now grounded the symbolic effort function:
\[
\text{effort}: R \subset \Sigma^* \to \mathbb{R}^+ \quad \text{// effort is applied to symbolic forms that function as rules in structural logic}
\]

The three effort metrics are summarized as follows:

\begin{table}[h!]
\centering
\begin{tabular}{|l|c|l|}
\hline
\textbf{Metric} & \textbf{Formula} & \textbf{Meaning} \\
\hline
Frequency-based & $-\log_2(N(T)/Z)$ & Commonness of use \\
Entropy-based   & $-\sum p_T \log_2 p_T$ & Path diversity participation \\
Drift-based     & $\Delta\partial\chi$ impact per latency unit & Coherence-preserving ability \\
\hline
\end{tabular}
\caption{Summary of symbolic effort metrics}
\end{table}

\medskip

Combined, these metrics form a principled foundation for symbolic latency, entropy, and structural thermodynamics.
