\chapter{Statistical RuleEvolution} \label{chapter-3-statistical-ruleevolution}

In Volume I, the RuleEvolution operator $\mathcal{R}E$ was defined as a coherence-based symbolic filtering process over $\Sigma^*$. It adjusts which symbolic structures remain in $R \subset \Sigma^*$, based on return behavior and structural coherence.

While this deterministic model suffices for foundational identity logic, it does not scale to complex symbolic systems with divergent reuse patterns and entropic fluctuations.

\medskip

In this chapter, we extend RuleEvolution into a statistical framework. This allows the system to evaluate transformations probabilistically, using symbolic entropy, reuse load, and coherence pressure to adjust transformation likelihoods across the Rule Set.

\medskip

We define a coherence score distribution, use it to compute update probabilities for each transformation $T \in R$, and then define the statistical RuleEvolution operator:
\begin{equation} \label{eq:ruleevolution-stat}
\mathcal{R}E_{\text{stat}} : (R_t, \chi_t, H_t, P_t) \longrightarrow \mathbb{P}_{t+1}(R)
\end{equation}

This enables stochastic symbolic evolution and supports thermodynamic analogs in later chapters.

\section{Coherence Score Distribution} \label{coherence-score-distribution}

Let $T \in R$, and let $[A] \in \Omega_3$ be any identity such that $T \in \mathcal{L}([A])$.

\begin{definition}[Coherence Score] \label{def:coherence-score}
The coherence score of $T$ is defined as:
\begin{equation} \label{eq:coherence-score}
S(T) := \sum_{[A] \in \mathcal{U}(T)} \left( \Delta\partial\chi([A], T) + \Delta\rho([A], T) \right)
\end{equation}
\end{definition}

Where:
\begin{itemize}
    \item $\mathcal{U}(T)$ is the set of all identities influenced by $T$,
    \item $\Delta\partial\chi([A], T)$ is the change in signature drift caused by $T$,
    \item $\Delta\rho([A], T)$ is the change in collapse robustness caused by $T$.
\end{itemize}

\paragraph{Interpretation.}
$S(T)$ is treated as a symbolic fitness metric: it quantifies the contribution of $T$ to structural coherence. Higher scores indicate more coherence-stabilizing transformations.

\section{Probabilistic Retention and Removal} \label{probabilistic-retention-and-removal}

Let $\theta_a$, $\theta_d$ be system parameters representing promotion and demotion sensitivity, respectively.

\begin{definition}[Update Probabilities]
We define the probability of retaining and promoting transformation $T$ based on its coherence score $S(T)$:
\begin{align}
\mathbb{P}_{\text{retain}}(T)  &:= \sigma\left(\frac{S(T) - \theta_d}{\delta}\right) \label{eq:retain-probability} \\
\mathbb{P}_{\text{promote}}(T) &:= \sigma\left(\frac{S(T) - \theta_a}{\delta}\right) \label{eq:promote-probability}
\end{align}
\end{definition}

Here, $\sigma$ is the logistic function defined as:
\begin{equation} \label{eq:logistic-function}
\sigma(x) := \frac{1}{1 + e^{-x}}
\end{equation}
and $\delta$ is a sharpness parameter controlling the slope of the transition.

\paragraph{Interpretation.}
\begin{itemize}
    \item If $S(T) \gg \theta_a$, $T$ is promoted with high probability.
    \item If $S(T) \ll \theta_d$, $T$ is removed with high probability.
    \item For intermediate values, the update is soft and stochastic.
\end{itemize}

\section{Rule Set Distribution Update} \label{rule-set-distribution-update}

Let $\mathbb{P}_t(R)$ be the probability distribution over the Rule Set $R$ at time $t$.

\begin{definition}[Rule Update Equation]
The update rule for transformation $T \in R$ is given by:
\begin{equation} \label{eq:rule-update}
\mathbb{P}_{t+1}(T) := \lambda \cdot \mathbb{P}_t(T) + (1 - \lambda) \cdot f(S(T))
\end{equation}
\end{definition}

where:
\begin{itemize}
    \item $\lambda \in [0, 1]$ is a retention weight,
    \item $f(S(T))$ is the normalized fitness score computed as:
    \begin{equation} \label{eq:fitness-softmax}
    f(S(T)) := \frac{\exp(S(T))}{\sum_{T' \in R} \exp(S(T'))}
    \end{equation}
\end{itemize}

This defines a softmax-weighted evolutionary distribution over the rule set, balancing persistence and adaptation.

\section{Statistical RuleEvolution Operator} \label{statistical-ruleevolution-operator}

We now define the statistical RuleEvolution operator as:
\begin{equation} \label{eq:RE-stat-operator}
\mathcal{R}E_{\text{stat}} : (R_t, \chi_t, H_t, P_t) \longrightarrow \mathbb{P}_{t+1}(R)
\end{equation}

\paragraph{Inputs.}
\begin{itemize}
    \item $R_t \subset \Sigma^*$: the current set of coherence-permitting symbolic structures (rules),
    \item $\chi_t$: coherence signature field at time $t$,
    \item $H_t$: entropy field over $\Omega_3$,
    \item $P_t$: coherence pressure map.
\end{itemize}

\paragraph{Outputs.}
$\mathcal{R}E_{\text{stat}}$ computes:
\begin{itemize}
    \item Coherence scores $S(T)$ for all $T \in R_t$,
    \item Promotion and removal probabilities,
    \item Updated symbolic distribution $\mathbb{P}_{t+1}(R)$ over the Rule Set.
\end{itemize}

\section{Transition to Deterministic Limit} \label{transition-to-deterministic-limit}

If $\lambda \to 0$ and $\delta \to 0$, the system behavior becomes deterministic:

\begin{itemize}
    \item $T$ is promoted if $S(T) > \theta_a$,
    \item $T$ is removed if $S(T) < \theta_d$.
\end{itemize}

In this limit, $\mathcal{R}E_{\text{stat}}$ reduces to the original deterministic RuleEvolution operator defined in Volume I.

\medskip

Thus, the deterministic $\mathcal{R}E$ is a special case of the statistical operator, emerging when stochastic modulation vanishes.

\section{Summary} \label{summary-1}

The statistical RuleEvolution operator defines symbolic transformation update as a probability distribution evolving under coherence pressure.

\begin{table}[h!]
\centering
\begin{tabular}{|l|c|l|}
\hline
\textbf{Component} & \textbf{Symbol} & \textbf{Role} \\
\hline
Coherence Score         & $S(T)$                  & Fitness signal for rule utility       \\
Promotion Probability   & $\mathbb{P}_{\text{promote}}(T)$ & Soft addition rate                  \\
Retention Probability   & $\mathbb{P}_{\text{retain}}(T)$  & Soft survival rate                  \\
Update Rule             & $\mathbb{P}_{t+1}(T)$   & Distribution shift                   \\
Evolution Operator      & $\mathcal{R}E_{\text{stat}}$     & Governs symbolic adaptation         \\
\hline
\end{tabular}
\caption{Components of the Statistical RuleEvolution framework}
\end{table}

\medskip

This prepares the formal structure for coherence thermodynamics and symbolic system evolution in the presence of entropy and reuse fluctuations.
