
\appendix
\section*{Appendix B: Derivation of the Koide Formula from Euler-Coherent Triplets} \label{appendix:koide-proof}
\addcontentsline{toc}{section}{Appendix B: Derivation of the Koide Formula from Euler-Coherent Triplets}

\subsection*{B.1 Euler-Coherent Triplet Structure}

We begin with the structural foundation established in Volume VIII: the coherence of a stable triplet identity arises from balanced return symmetry over Euler's unit circle.

\begin{postulate}[Phase-Balanced Coherence]
Let $\theta_i \in \mathbb{R}$ be the symbolic phase angle of identity $[A_i]$ on the complex unit circle. A coherence-locked triplet satisfies:
\[
\sum_{i=1}^3 e^{i\theta_i} = 0
\]
\end{postulate}

This condition corresponds to three points equally spaced around the unit circle. Without loss of generality, we choose:
\[
\theta_0 = 0,\quad \theta_1 = \frac{2\pi}{3},\quad \theta_2 = \frac{4\pi}{3}
\]

The three coherence states are thus:
\[
z_0 = 1,\quad z_1 = e^{2\pi i / 3},\quad z_2 = e^{4\pi i / 3}
\]
which satisfy:
\[
z_0 + z_1 + z_2 = 0
\]

This geometric structure is what we refer to as a \textit{Euler-coherent triplet}.

\subsection*{B.2 Mass as Projection from Phase Angle}

In the SCM framework, mass is not a primitive quantity—it emerges from the structure of return. For a coherence identity $[A_i]$ with symbolic phase angle $\theta_i$, we define its projected return effort (interpreted as root mass) as:

\begin{definition}[Phase-Projected Mass]
Let $\sqrt{m_i} := a + b \cos(\theta_i)$ for real constants $a, b$. Then the mass of identity $[A_i]$ is given by:
\[
m_i := (a + b \cos(\theta_i))^2
\]
\end{definition}

This form:
\begin{itemize}
    \item Is symmetric under phase rotation,
    \item Reduces to coherence magnitude when $a = 0$,
    \item Embeds directly in Euler geometry, since $\cos \theta_i$ is the real part of $e^{i\theta_i}$.
\end{itemize}

Using:
\[
\theta_0 = 0,\quad \theta_1 = \frac{2\pi}{3},\quad \theta_2 = \frac{4\pi}{3}
\Rightarrow
\cos \theta_0 = 1,\quad \cos \theta_1 = \cos \theta_2 = -\frac{1}{2}
\]

We find:
\begin{align*}
\sqrt{m_0} &= a + b \\
\sqrt{m_1} &= \sqrt{m_2} = a - \frac{b}{2}
\end{align*}
\[
m_0 = (a + b)^2,\quad m_1 = m_2 = \left( a - \frac{b}{2} \right)^2
\]

\subsection*{B.3 Contradiction Functional and Normalization}

\begin{definition}[Contradiction Functional]
Let the normalized root mass deviation from the mean be:
\[
X_\pi([A_i]) := \frac{\sqrt{m_i} - \bar{s}}{\bar{s}},\quad \text{where } \bar{s} := \frac{1}{3} \sum_{i=0}^2 \sqrt{m_i}
\]
Then the total contradiction is:
\[
\mathcal{C}(a, b) := \sum_{i=0}^2 X_\pi([A_i])^2
\]
\end{definition}

Since:
\[
\bar{s} = \frac{1}{3} \left( a + b + 2(a - \frac{b}{2}) \right) = a
\]

We have:
\[
X_\pi([A_0]) = \frac{b}{a},\quad
X_\pi([A_1]) = X_\pi([A_2]) = -\frac{b}{2a}
\]

Then:
\[
\mathcal{C}(a, b) = \left( \frac{b}{a} \right)^2 + 2 \left( \frac{b}{2a} \right)^2 = \frac{3b^2}{2a^2}
\]

\subsection*{B.4 Apply Koide Functional and Solve}

Define:
\[
K(a, b) := \frac{m_0 + m_1 + m_2}{\left( \sqrt{m_0} + \sqrt{m_1} + \sqrt{m_2} \right)^2}
\]

Compute numerator:
\[
m_0 + m_1 + m_2 = (a + b)^2 + 2 \left( a - \frac{b}{2} \right)^2 = 3a^2 + \frac{3b^2}{2}
\]

Compute denominator:
\[
\left( \sum \sqrt{m_i} \right)^2 = (a + b + 2a - b)^2 = (3a)^2 = 9a^2
\]

Then:
\[
K(a, b) = \frac{3a^2 + \frac{3b^2}{2}}{9a^2} = \frac{1}{3} + \frac{b^2}{6a^2}
\]

Let $x := \frac{b^2}{a^2}$, then:
\[
K = \frac{1}{3} + \frac{x}{6},\quad
\mathcal{C} = \frac{3}{2}x \Rightarrow
x = 6K - 2,\quad
\mathcal{C}(K) = 9K - 3
\]

\paragraph{Structural Constraint.}
Although $\mathcal{C}(a, b)$ achieves a minimum at $b = 0$, this corresponds to equal mass: $\sqrt{m_0} = \sqrt{m_1} = \sqrt{m_2}$. Such a state:
\begin{itemize}
    \item Is degenerate and structurally forbidden under SCM reuse logic,
    \item Violates RuleEvolution separation criteria (Volume IV, Chapter 25),
    \item Triggers collapse due to indistinguishability and symmetry lock.
\end{itemize}

\paragraph{Conclusion.}
The lowest contradiction state under phase-locked reuse, excluding degenerate collapse, is:
\[
\boxed{
K = \frac{2}{3},\quad \mathcal{C} = 3
}
\]
This completes the derivation: the Koide ratio emerges as the unique, non-degenerate minimum contradiction state in Euler-coherent χ-space.
