\chapter{The Eulerian Coherence Singularity} \label{chapter:eulerian-singularity}

In classical cosmology, the Big Bang is described as a physical singularity—an initial state of infinite density and zero volume. In the SCM framework, this interpretation is replaced by a structural one. The origin of the universe is the **first non-annihilating, coherence-closed return loop**.

\section{Pre-Identity: Ω₋ and Structural Cancellation}

Before the emergence of persistence, all symbolic activity resolves into annihilation. Each transformation satisfies:

\[
[A] \to [-A] \to [0], \quad \text{with } X_\pi = -1
\]

This domain is defined as Ω₋, the annihilation zone of coherence inversion. Any symbolic structure that resolves to $[0]$ through inversion belongs to Ω₋. There is no persistence, no reuse, no phase continuity.

At this stage, the coherence topology is **empty**:
- No identity loops exist.
- No symbolic reuse is permitted.
- All return paths reduce contradiction through immediate cancellation.

\section{First Identity as Euler Closure}

Let $[A]$ be the first identity for which a coherence-closed return loop exists such that:

\[
\sum_{i} \theta_i = 2\pi, \quad \text{with } \theta_i \in \mathbb{R},\ \theta_i < \pi
\]

\begin{definition}[Eulerian Coherence Singularity]
The Eulerian Singularity is the first structure $[A]$ for which:

\begin{itemize}
  \item $\mathcal{R}([A]) \neq \emptyset$
  \item $\Psi([A]) := \sum e^{i\theta_i} = 1$
  \item $\Lambda([A]) = \Lambda_0$
  \item $X_\pi([A]) = 0$, \quad $X_\epsilon([A]) = 0$
\end{itemize}

It is the first identity satisfying return under coherence without collapse.
\end{definition}

This corresponds to the first **non-inverting**, **coherence-locked** identity on the unit circle. It lies on the phase manifold as a closed loop of angle $2\pi$ and represents the **birth of persistence**.

\section{Symbolic Origin, Not Spacetime Origin}

The Eulerian singularity is not a location in spacetime. It is the **first resolution anchor in the phase topology of return**.

\begin{itemize}
  \item There is no metric geometry.
  \item There is no entropy—return sets are minimal and coherent.
  \item The only structure is a symmetric 2-step return loop $[A] \to [B] \to [A]$, with coherence-preserving transformation and no inversion: $[B] \ne [-A]$ and $[A] + [B] \ne [0]$
\end{itemize}

This structure is what later volumes identify as the **photon**:
- Minimal latency: $\Lambda = \Lambda_0$
- Zero drift: $\partial \chi = 0$
- Perfect return symmetry: $X_\pi = 0$

\section{The Breaking of Ω₋ Symmetry}

Persistence requires an escape from Ω₋. This occurs when a symbolic structure avoids cancellation by completing its return path without inversion. Formally:

\[
[A] + [B] = [A], \quad \text{with } \Psi([A]) = 1
\]

This coherence escape defines the **start of structure** in the SCM universe.

\section{Global Consequences}

The Eulerian coherence singularity defines:

\begin{itemize}
  \item The **first identity** capable of reuse ($X_\epsilon = 0$)
  \item The **start of symbolic inflation** (Chapter 2)
  \item The **first nonzero value** of the mass threshold: $\Lambda_0 > 0$
  \item The **minimum phase requirement** for persistent structure
\end{itemize}

\begin{theorem}[Minimum Return Angle]
Any identity $[A]$ such that $\sum \theta_i < 2\pi$ cannot persist under RuleEvolution. Therefore, $2\pi$ is the minimal coherence angle, and $\Lambda_0$ is the minimal symbolic effort for persistent existence.
\end{theorem}

This is the origin of **mass, time, and structure**. The Eulerian singularity is not a physical event, but a **topological emergence condition**.

---

\section{Conclusion}

The Big Bang is redefined as the **first Eulerian loop**. Not a burst of energy, but the first successful symbolic closure across a $2\pi$ return cycle.

From this point forward, RuleEvolution operates, coherence expands, and contradiction pressure builds. What follows is not a spacetime expansion—but an explosion of phase-locked return graphs.

In the next chapter, we describe this expansion as **symbolic inflation** driven by structural freedom and reuse creation.
