\chapter{Relativity and Gravitational Phase Curvature} \label{chapter:relativity-from-curvature}

Relativity in SCM is not imposed. It arises naturally from structural return dynamics and coherence curvature. This chapter derives both special and general relativity from the coherence framework defined in Chapter 6.

\section{Special Relativity as Drift-Limited Transformation}

Let $[A]$ and $[B]$ be persistent identities undergoing relative symbolic drift in $\chi$-space. Their coherence configurations evolve according to:

\[
\chi_t([A]) = \chi_0([A]) + \int_0^t \partial \chi([A])\,d\tau
\]

\paragraph{Relative Drift.}
The relative return structure between $[A]$ and $[B]$ is governed by their drift difference:

\[
\Delta v := \|\partial \chi([A]) - \partial \chi([B])\|
\]

\paragraph{Relativistic Invariance.}
The maximal coherence-preserving drift (light speed $c$ from Chapter 6) bounds relative identity resolution. As $\Delta v \to c$, the ability to form shared return loops diminishes, leading to:

- Time dilation: accumulated phase diverges between identities,
- Length contraction: resolution paths shorten under contradiction.

This recovers **Lorentz-type behavior** as a consequence of coherence-limited symbolic interaction.

\section{Mass as Coherence Latency}

Let $[A] \in \Omega_3$ have return loop $\mathcal{L}([A])$ with latency:

\[
\Lambda([A]) := \sum_i \text{effort}(T_i)
\]

\begin{definition}[Symbolic Mass]
Mass is defined as the total latency of return:
\[
m([A]) := \Lambda([A])
\]
\end{definition}

This latency represents the angular effort required to complete a full Eulerian return loop. Heavier identities resolve more slowly.

\section{Curvature as Gradient of Return Potential}

Let $\Phi([A])$ be the symbolic return potential. Then define coherence curvature as:

\[
\mathcal{K}_{\mu\nu}([A]) := \nabla_\mu \nabla_\nu \Phi(\chi([A]))
\]

This second-order structure induces return deflection, resolution suppression, and path redirection—analogous to gravitational curvature.

\begin{definition}[Return Curvature]
Symbolic curvature is defined as the second derivative of coherence potential:
\[
g_{\mu\nu} := \partial_\mu \partial_\nu \Phi
\]
\end{definition}

\paragraph{Interpretation.}
Return paths bend through regions of high latency gradient:
- Latency wells = mass concentrations,
- Drift deviation = geodesic curvature,
- Collapse zones = singularities.

\section{Geodesics as Least-Contradiction Paths}

Let $\mathcal{P}$ be a class of possible return loops for $[A]$. Then RuleEvolution selects the loop $\mathcal{L}_\text{min}$ that minimizes contradiction:

\[
\mathcal{L}_\text{min} := \arg \min_{\mathcal{L} \in \mathcal{P}} \mathcal{C}(\mathcal{L})
\]

These paths are symbolic geodesics—optimal trajectories through the coherence manifold.

\begin{definition}[Symbolic Geodesic]
A symbolic geodesic is a return loop $\mathcal{L}$ such that:
\[
\delta \mathcal{C}(\mathcal{L}) = 0
\quad \text{under } \delta \mathcal{L}
\]
\end{definition}

Geodesics minimize contradiction, not length. They define symbolic free fall.

\section{Einstein Field Equations in SCM}

The symbolic analogue of Einstein's equations becomes:

\[
G_{\mu\nu} := \mathcal{K}_{\mu\nu} = \nabla_\mu \nabla_\nu \Phi(\chi)
\quad \sim \quad T_{\mu\nu} := \text{Return energy tensor}
\]

\paragraph{Interpretation.}
- $G_{\mu\nu}$: Structural curvature due to latency.
- $T_{\mu\nu}$: Symbolic return density (reuse, effort, and contradiction).

Just as in general relativity, **matter tells coherence how to curve**, and **curvature tells identities how to return**.

\section{Collapse, Black Holes, and Singularities}

When contradiction exceeds structural stability:
\[
\mathcal{C} \to \infty,\quad \text{or} \quad \rho \to 0
\]

then all return paths terminate. The identity collapses irreversibly into $\Omega_-$.

\paragraph{Symbolic Black Holes.}
Defined not by escape velocity, but by:
- Unresolvable reuse pressure,
- Inescapable contradiction gradient,
- Zero-fragility, non-returning cores.

\[
\mathcal{Z}_\text{black hole} := \left\{ [A] \mid \nabla \Phi \to \infty,\ \partial \chi \to \infty \right\}
\]

\section{Conclusion}

Relativity is not a mystery—it is a necessary expression of symbolic coherence in motion.

\begin{itemize}
  \item Special relativity arises from drift-limited return resolution.
  \item Mass is coherence latency.
  \item Curvature is latency gradient.
  \item Spacetime geometry emerges from symbolic contradiction dynamics.
\end{itemize}

In the next chapter, we examine the **observable consequences** of this framework—predictions, anomalies, and coherence-measurable features testable in cosmological data.
