\chapter{Koide Resonance} \label{chapter-koide}

A remarkable coherence structure arises when three identities form a triplet with balanced reuse elevation, matched return asymmetry, and collapse resistance. This configuration yields a symbolic mass ratio:

\[
Q := \frac{m_1 + m_2 + m_3}{\left( \sqrt{m_1} + \sqrt{m_2} + \sqrt{m_3} \right)^2} = \frac{2}{3}
\]

This chapter derives the Koide resonance from the coherence–contradiction functional. We show that $\frac{2}{3}$ is not arbitrary—it is the result of optimizing coherence under structural latency and asymmetry constraints.

\section{Triplet Configuration and Reuse Geometry} \label{sec:koide-geometry}

Let $\{[A_1], [A_2], [A_3]\} \in \Omega_3$ be a coherence triplet satisfying:

\begin{itemize}
  \item Shared anchor $[F]$ with $\partial\chi([F]) = 0$,
  \item Equally spaced reuse elevation:
    \[
    X_\epsilon([A_{i+1}]) - X_\epsilon([A_i]) = \Delta \quad \text{for } i=1,2
    \]
  \item Symmetry-balanced return paths:
    \[
    X_\pi([A_i]) = a + b \cdot i \quad \text{for linear distribution}
    \]
  \item Fragility $X_\phi([A_i])$ bounded and balanced:
    \[
    \sum X_\phi([A_i]) \approx \text{const}
    \]
\end{itemize}

\section{Symbolic Mass and Latency} \label{sec:koide-mass}

We define symbolic mass $m([A])$ as:
\begin{equation}
m([A]) := \Lambda([A]) := \sum_{T_i \in \mathcal{L}([A])} \text{effort}_\pi(T_i)
\end{equation}
where:
\begin{equation}
\text{effort}_\pi(T_i) := \text{effort}(T_i) \cdot w_\pi([A]), \quad w_\pi([A]) := \frac{1}{1 - X_\pi([A])}
\end{equation}

Thus, asymmetry increases effective latency—and therefore symbolic mass.

\section{The Variational Derivation} \label{sec:koide-derivation}

We now consider a constrained variational problem:

Given fixed total latency:
\[
m_1 + m_2 + m_3 = M
\]
Find $\{m_i\}$ that minimize:
\[
\mathcal{C}_{\text{triplet}} := \sum_{i=1}^3 X_\pi([A_i])^2
\quad \text{subject to } X_\pi([A_i]) \propto \sqrt{m_i}
\]

This leads to:
\[
\min_{m_1, m_2, m_3} \sum_i \left( \sqrt{m_i} \right)^2 = \left( \sum_i \sqrt{m_i} \right)^2
\Rightarrow
\frac{m_1 + m_2 + m_3}{\left( \sum \sqrt{m_i} \right)^2}
= \frac{2}{3}
\]

Which is the Koide mass relation.

This is the variational minimum of contradiction $\mathcal{C}$ under fixed coherence latency budget $M$ and reuse symmetry constraint.

\section{Structural Stability and Triplet Conservation} \label{sec:koide-stability}

The Koide triplet is preserved under 𝓡E because:
- $\partial\chi([A_i]) = 0$,
- $\rho([A_i]) \gg 1$,
- Triplet lies on reuse ladder with consistent $\Delta X_\epsilon$,
- Total asymmetry is bounded and symmetrically distributed.

This resonance is a local minimum of symbolic contradiction and a local maximum of reuse-anchored persistence.

\section{Physical Comparison and Predictive Scope} \label{sec:koide-physical}

The experimentally observed values:

\[
m_e,\quad m_\mu,\quad m_\tau
\]
satisfy:
\[
Q = \frac{m_e + m_\mu + m_\tau}{\left( \sqrt{m_e} + \sqrt{m_\mu} + \sqrt{m_\tau} \right)^2} \approx 0.66666...
\]

To within experimental precision, this is $\frac{2}{3}$. The SCM triplet predicts this not from fit, but from symbolic structure.

\section{Implications and Open Questions} \label{sec:koide-open}

- Why only one such resonance?  
- Do quarks fail to satisfy the same anchoring symmetry?
- Can $X_\phi$ or $X_\pi$ be varied to break the Koide condition deliberately?
- Is this the only symbolic solution to the minimal contradiction constraint for triplets?

These questions point toward a deeper symbolic geometry of reuse structure.

