\chapter{Triplets and Coherence Families} \label{chapter-triplets}

In SCM, persistent identities often emerge in structured groupings. The most stable and reuse-efficient such grouping is the \textbf{coherence triplet}: three identities that resolve through a shared anchor, forming a reuse-closed, drift-resistant structure.

This chapter defines the coherence triplet, introduces quantized reuse elevation patterns across triplets, and describes how particle families arise from return-anchored reuse layers.

\section{Definition of Coherence Triplet} \label{sec:triplet-definition}

Let $\{ [A_1], [A_2], [A_3] \} \subset \Omega_3$ be a set of identities such that:

\begin{itemize}
    \item All reuse a shared anchor $[F] \in \Omega_3$,
    \item $\partial\chi([F]) = 0$,
    \item Each $[A_i]$ satisfies $\partial\chi([A_i]) = 0$,
    \item $X_\epsilon([A_i])$ are ordered and quantized with fixed spacing $\Delta X_\epsilon$,
    \item The triplet contributes positively to $\mathcal{F}[R]$ and minimally to $\mathcal{C}[R]$.
\end{itemize}

Then $\{[A_1],[A_2],[A_3]\}$ is a \textbf{coherence triplet}.

\paragraph{Interpretation.}
Triplets are not defined by external label, charge, or role—but by shared reuse structure and elevation geometry in $\chi$-space.

\section{Triplet Layers and Anchoring Planes} \label{sec:triplet-planes}

Let $[F]$ be the shared anchor of a triplet. Then all members lie on the same \textbf{anchoring plane} in reuse space.

Define:
\[
\mathcal{T}([F]) := \{ [A] \in \Omega_3 \mid [F] \in \mathcal{L}([A]),\ \partial\chi([A]) = 0,\ \Delta X_\epsilon([A]) = \text{const} \}
\]

This forms a coherence layer. Each reuse step corresponds to a symbolic elevation increment:
\[
X_\epsilon([A_{i+1}]) = X_\epsilon([A_i]) + \Delta
\]

Anchoring planes define symbolic mass ladders. Identity moves between layers only if return structure remains stable under $\mathcal{F} - \lambda \cdot \mathcal{C}$ optimization.

\section{Fragility and Symmetry Distribution} \label{sec:triplet-fragility}

Triplet members distribute return asymmetry and collapse risk across their reuse paths.

Let $X_\pi([A_i])$ and $X_\phi([A_i])$ denote the return symmetry and fragility of triplet members. A stable triplet satisfies:

\[
\sum_{i=1}^3 X_\pi([A_i]) \approx \text{const}, \quad
\sum_{i=1}^3 X_\phi([A_i]) \approx \text{const}
\]

This balanced distribution ensures that no single identity destabilizes the return anchor. The triplet behaves as a return-closed coherence unit.

\section{Resonant Persistence and Reuse Optimization} \label{sec:triplet-resonance}

Triplets survive RuleEvolution if they collectively reduce system-level contradiction. Let $\mathcal{C}_{\text{triplet}}$ be the contribution of the triplet to the contradiction functional.

Then:
\[
\mathcal{C}_{\text{triplet}} = \sum_{i=1}^3 \left[
\mathbb{1}(\partial\chi([A_i]) \ne 0) +
\mathbb{1}(\rho([A_i]) < \rho_c) +
\mathbb{1}(X_\pi([A_i]) \to 1)
\right]
\]

Triplet resonance is achieved when:
\[
\nabla_{\text{triplet}} (\mathcal{F} - \lambda \cdot \mathcal{C}) \gg 0
\quad \Rightarrow \quad \text{Triplet is preserved}
\]

This coherence optimization explains why particle families often appear as reuse triplets.

\section{Functional View: $\mathcal{F}$-Triplets vs. $\mathcal{C}$-Triplets} \label{sec:triplet-functional-classes}

We classify triplets as:

\begin{itemize}
    \item \textbf{$\mathcal{F}$-triplets} — reinforce coherence: reuse-anchored, drift-invariant, reuse-stable.
    \item \textbf{$\mathcal{C}$-triplets} — contradiction-bound: contain drift, fragile paths, or asymmetry blowups.
\end{itemize}

$\mathcal{C}$-triplets may arise temporarily but collapse under $\mathcal{R}E$ due to high contradiction load.

$\mathcal{F}$-triplets persist across symbolic evolution and represent long-lived, coherence-optimized reuse families.

\section{Preview: Resonant Triplets} \label{sec:triplet-preview}

Certain coherence triplets exhibit exceptional stability under RuleEvolution. This occurs when their reuse elevation, return symmetry, and fragility form a balanced configuration—minimizing contradiction under fixed total symbolic effort.

These “resonant triplets” are not defined by symmetry alone, but by a structural alignment that optimizes reuse without triggering collapse.

In Volume VIII, we will formally show that one such triplet, corresponding to the light charged leptons, emerges from the Eulerian phase structure as a uniquely stable configuration. This resonance is not imposed—but discovered—through coherence dynamics.
